CODECHECK: An Open Science initiative for the independent execution of computations underlying research articles during peer review to improve reproducibility

### 2021-02-23 codecheck-paper.tex
spell-checked; removed a couple of abbrevs

We also see a great opportunity to involve \emph{early-career
  researchers} (ECRs) as codecheckers.  ECRs arguably have a high
interest in learning about new tools and technologies, to build up
their own expertise.  CODECHECK offers a way for ECRs to gain insights
into new research and highlight the importance of reproduction.
\emph{ReScience X}, a journal devoted to reproduction and replication
experiments \cite{roesch_new_2020}, shares an interest in this
combination.  ECRs are also often familiar with new technologies, thus
also making them likely to author CODECHECK-ready manuscripts.  A
supporting data point for ECRs as early adopters is that they are
responsible for 77\% of 141 registered reports that were submitted
\cite{chambers_registered_2019}.  As ECRs are introduced to peer
review as codecheckers, they may transition into the role of
peer-reviewer over time.  Overall, we see several opportunities
and benefits to setting up a new process for CODECHECKING with a clear
commitment to openness and transparency, independent of the current
peer review process (see \emph{Openness} dimension).

@@

\textbf{5. Open by default and transitional by disposition.} \\
Unless there are strong reasons to the contrary
(e.g.,~sensitive data on human subjects), all code and data, both from
author and codechecker, will be made freely available when
the certificate is published.  Openness is not required for the paper
itself, to accommodate journals in their transition to
Open Access models.  The code and data publication should follow
community good practices.  Ultimately we may find that CODECHECKS are
subsumed within peer review.
%% By disposition, a CODECHECK workflow should
%% be designed (a) recognising shortcomings in asserting reproducibility
%% of submissions and (b) with an understanding of the complex situation
%% that led to this. However, establishing a CODECHECK workflow should
%% mirror initiatives to improve education around reproducibility of
%% methods.  The long term goal is for a specific community to reach a
%% state where the level of verification ensured by codecheckers becomes
%% part of peer review and a regular merit of all accepted papers.

@@

The third theme represents co-author DN's service as a Reproducibility
Reviewer at the AGILE conference series, where the \emph{Reproducible
  AGILE} Initiative \cite{reproducible_agile} independently
established a process for reproducing workflows at the AGILE
conference series \cite{nust_improving_2020}.  While using slightly
different terms and infrastructure (``reproducibility reports'' are
published on the Open Science Framework (OSF) instead of certificates
on Zenodo) AGILE reproducibility reviews adhere to CODECHECK
principles.  A few checks were also completed as part of peer reviews
for GIScience journals.

@@

We use freely available infrastructure, GitHub and Zenodo, to run our
system.  The \texttt{codecheckers} \textbf{GitHub} organisation at
\url{https://github.com/codecheckers} contains projects for managing
the project website, the codecheckers community and its discussions,
code repositories, and the main register of CODECHECKs. Both the
project website \url{https://codecheck.org.uk/} and the register at
\url{https://codecheck.org.uk/register} are hosted as GitHub pages.
The register database is a single table in CSV format that connects
the certificate identifier with the repository associated with a
CODECHECK. Each of these repositories, which currently can be hosted
on GitHub or Open Science Framework (OSF), contains the CODECHECK
metadata file \texttt{codecheck.yml}. The register further contains a
column for the type of check, e.g.,~community, journal, or conference,
and the respective GitHub issue where communications and assignments
around a specific check are organised. No information is duplicated
between the register and the metadata files. The continuous
integration infrastructure of GitHub, GitHub Actions, is used to
automate the register for viewing (HTML) and for integration with
other services (CSV, JSON).  \textbf{Zenodo} is our preferred open
repository for storing certificates. It mints DOIs for deposits and
ensures long-term availability of all digital artefacts related to the
project. The CODECHECK community on Zenodo is available at
\url{https://zenodo.org/communities/codecheck/}. It holds
certificates, the regularly archived register
\cite{codecheck_register_jan2021}, and other material related to
CODECHECK.

@@

For research with high stakes, where reproduction would be too weak and
post-publication replication possibly too late because of policy impact,
Benjamin-Chung et~al. \cite{benjamin-chung_internal_2020} propose
\emph{internal replication}.
A workflow that has undergone internal
replication would likely be of high quality and relatively easy to check.
Similarly, internal CODECHECKs may be used, with the same limitations such
as group think \cite{benjamin-chung_internal_2020},
to ensure reproducibility before submission.
Such internal checks are professionalised in local reproduction services,
such as CISER R-squared
(\url{https://ciser.cornell.edu/research/results-reproduction-r-squared-service/})
or YARD (\url{https://isps.yale.edu/research/data/approach}),
or in communities such as Oxford's code review network (OxCRN,
\url{https://github.com/OxfordCodeReviewNet/forum}).

@@

Another platform for publishing results of reproductions is
\emph{SciGen.Report} (\url{https://scigen.report/}).  It is a
community-run independent platform to foster communication on
reproducibility.  People can report on fully, partially, or
un-successful reproductions of articles after publication.

@@

A codechecker does not fulfill the same role as a \emph{statistical
  reviewer}, as it is applied by some journals in the biomedical
domain (cf.~\cite{petrovecki_role_2009,greenwood_how_2015}).  The
statistical reviewer evaluates the appropriateness of statistical
methods \cite{greenwood_how_2015} and can support topical reviewers
if, e.g.,~complex methods or sophisticated variants of statistical
tests are applied \cite{petrovecki_role_2009}.  The codechecker may go
equally deep into the review, but only if they have the expertise and
time. We can imagine a tiered workflow where a codechecker could, just
as a conventional reviewer could, recommend a detailed code review
(see next paragraph) to the editor if they come upon certain issues while
examining the work.

@@

\textbf{Can't someone cheat?} Yes. We simply check that the code runs,
not that is correct or sound science. This ``mechanical'' test is
indeed a low bar.  By having code and data openly deposited, third
parties can later examine the code, and we hope that knowing the code
will be open ensures that authors will not cheat.  It also allows
researchers, potentially with new methods, to look for errors.  This
is more effective than engaging in an arms race on building methods to
detect malicious intent now with closed datasets and code.  This is
analogous to storing blood samples of sport champions today to
possibly detect doping in the future with more sensitive methods
(cf. \cite{everythinghertz97}).  Another comparison that helped us
define the scope of a CODECHECK is that we think of the codechecker as
forensic photographer, capturing details so that an investigator may
later scrutinize them.

\textbf{Who's got time for more peer review?} Agree; codechecking
takes time that could otherwise be used for traditional peer review.
However, a CODECHECK is different from peer review.  First, the
\textbf{technical nature} of a CODECHECK sets clear expectations and
thereby time budget compared to conventional peer review.  For
example, authors are told what to provide and the codechecker can be
told when to stop.  Codecheckers can always directly ask the author
when clarification is required, thereby increasing
\textbf{efficiency}.  Second, the specific skill set of a codechecker
allows for \textbf{different groups} to participate in the review
process.  ECRs might be attracted to learn more about recent methods,
peer review, and reproducibility practices.  Research Software
Engineers (RSE) %% , see \url{https://en.wikipedia.org/wiki/Research_software_engineering})
who might not regularly be involved in writing or reviewing papers
might be interested in increasing their connection with scholarly
practices.  An extra codechecker may simplify the matchmaking an
editor does when identifying suitable reviewers for a submission,
as technical and topical expertise can be provided by different
people.  Third, recall that CODECHECKs should always be publicly
available, unlike peer review reports.  With code and workflows, the
codechecker's feedback may directly impact and improve the author's
work.  The public certificates and contributions provide \textbf{peer
  recognition} for the codechecker.  Fourth, we found that focusing on
workflow mechanics and interacting with the author makes reproductions
\textbf{educational}. It also is a different role and, as such, could
be a welcome option for researchers to give back their time to the
community.

@@

CODECHECK works --- we have created a considerable number of
certificates to demonstrate it.  The creation of certificates and
interactions with authors and editors shaped the principles and the
workflow and also confirmed the approach taken. This result
corroborates findings from similar evaluations of reproducible
computational research in journals and conferences.  CODECHECKs
increase transparency of the checked papers and can contribute to
building trust in scientific results.
% "science earns trust through transparency" > https://twitter.com/briannosek/status/1290240244101521408
The set of shared principles and common name, through recognition
value, will allow researchers to judge the level of scrutiny that
results have faced. CODECHECK requires direct acknowledgement of the
codechecker's contributions, not indirectly via citations of
reproductions or informal crediation and closed inner circles.

@@

A CODECHECK system, even if temporarily adopted as a sustainable
transition towards more open publication and review practices, can
contribute to increased trust in research outputs. Introducing
CODECHECK should be informed by lessons learned from (introducing)
open peer review (OPR) \cite{ross-hellauer_guidelines_2019}.  \textbf{Our
conversations with publishers and editors included procedural and
technical implications of openness, be it that OPR is neither applied
by nor familiar to stakeholders, or that process management software
would not support such open communication. WHAT DOES THIS SENTENCE
MEAN??  JUST DELETE?}

@@

% Local Variables:
% TeX-engine: xetex
% End:

>>>>

We also see a great opportunity to involve \emph{early-career
  researchers} (ECRs) as codecheckers.  ECRs arguably have a high
interest in learning about new tools and technologies, to build up
their own expertise.  CODECHECK offers a way for ECRs to gain insights
into new research and highlight the importance of reproduction.
\emph{ReScience X}, a journal devoted to reproduction and replication
experiments \cite{roesch_new_2020}, shares an interest in this
combination.  ECRs are also often familiar with new technologies, thus
also making them likely to author CODECHECK-ready manuscripts.  A
supporting data point for ECRs as early adopters is that they are
responsible for 77\% of 141 registered reports that were submitted
\cite{chambers_registered_2019}.  As ECRs are introduced to peer
review as codecheckers, they may transition into the role of
peer-reviewer over time.  Overall, we see several opportunities
and benefits to setting up a new process for codechecking with a clear
commitment to openness and transparency, independent of the current
peer review process (see \emph{Openness} dimension).

@@

\textbf{5. Open by default and transitional by disposition.} \\
Unless there are strong reasons to the contrary
(e.g.,~sensitive data on human subjects), all code and data, both from
author and codechecker, will be made freely available when
the certificate is published.  Openness is not required for the paper
itself, to accommodate journals in their transition to
Open Access models.  The code and data publication should follow
community good practices.  Ultimately we may find that CODECHECK activities are
subsumed within peer review.
%% By disposition, a CODECHECK workflow should
%% be designed (a) recognising shortcomings in asserting reproducibility
%% of submissions and (b) with an understanding of the complex situation
%% that led to this. However, establishing a CODECHECK workflow should
%% mirror initiatives to improve education around reproducibility of
%% methods.  The long term goal is for a specific community to reach a
%% state where the level of verification ensured by codecheckers becomes
%% part of peer review and a regular merit of all accepted papers.

@@

The third theme represents co-author DN's service as a Reproducibility
Reviewer at the AGILE conference series, where the \emph{Reproducible
  AGILE} Initiative \cite{reproducible_agile} independently
established a process for reproducing workflows at the AGILE
conference series \cite{nust_improving_2020}.  While using slightly
different terms and infrastructure (``reproducibility reports'' are
published on the Open Science Framework instead of certificates
on Zenodo) AGILE reproducibility reviews adhere to CODECHECK
principles.  A few checks were also completed as part of peer reviews
for GIScience journals.

@@

We use freely available infrastructure, GitHub and Zenodo, to run our
system.  The \texttt{codecheckers} \textbf{GitHub} organisation at
\url{https://github.com/codecheckers} contains projects for managing
the project website, the codecheckers community and its discussions,
code repositories, and the main register of CODECHECKs. Both the
project website \url{https://codecheck.org.uk/} and the register at
\url{https://codecheck.org.uk/register} are hosted as GitHub pages.
The register database is a single table in CSV format that connects
the certificate identifier with the repository associated with a
CODECHECK. Each of these repositories, which currently can be hosted
on GitHub or Open Science Framework, contains the CODECHECK
metadata file \texttt{codecheck.yml}. The register further contains a
column for the type of check, e.g.,~community, journal, or conference,
and the respective GitHub issue where communications and assignments
around a specific check are organised. No information is duplicated
between the register and the metadata files. The continuous
integration infrastructure of GitHub, GitHub Actions, is used to
automate the register for viewing (HTML) and for integration with
other services (CSV, JSON).  \textbf{Zenodo} is our preferred open
repository for storing certificates. It mints DOIs for deposits and
ensures long-term availability of all digital artefacts related to the
project. The CODECHECK community on Zenodo is available at
\url{https://zenodo.org/communities/codecheck/}. It holds
certificates, the regularly archived register
\cite{codecheck_register_jan2021}, and other material related to
CODECHECK.

@@

For research with high stakes, where reproduction would be too weak and
post-publication replication possibly too late because of policy impact,
Benjamin-Chung et~al. \cite{benjamin-chung_internal_2020} propose
\emph{internal replication}.
A workflow that has undergone internal
replication would likely be of high quality and relatively easy to check.
Similarly, internal CODECHECKs may be used, with the same limitations such
as group think \cite{benjamin-chung_internal_2020},
to ensure reproducibility before submission.
Such internal checks are professionalised in local reproduction services,
such as CISER R-squared
(\url{https://ciser.cornell.edu/research/results-reproduction-r-squared-service/})
or YARD (\url{https://isps.yale.edu/research/data/approach}),
or in communities such as Oxford's code review network (\url{https://github.com/OxfordCodeReviewNet/forum}).

@@

Another platform for publishing results of reproductions is
\emph{SciGen.Report} (\url{https://scigen.report/}).  It is a
community-run independent platform to foster communication on
reproducibility.  People can report on fully, partially, or
failed reproductions of articles after publication.

@@

A codechecker does not fulfil the same role as a \emph{statistical
  reviewer}, as it is applied by some journals in the biomedical
domain (cf.~\cite{petrovecki_role_2009,greenwood_how_2015}).  The
statistical reviewer evaluates the appropriateness of statistical
methods \cite{greenwood_how_2015} and can support topical reviewers
if, e.g.,~complex methods or sophisticated variants of statistical
tests are applied \cite{petrovecki_role_2009}.  The codechecker may go
equally deep into the review, but only if they have the expertise and
time. We can imagine a tiered workflow where a codechecker could, just
as a conventional reviewer could, recommend a detailed code review
(see next paragraph) to the editor if they come upon certain issues while
examining the work.

@@

\textbf{Can't someone cheat?} Yes. We simply check that the code runs,
not that is correct or sound science. This ``mechanical'' test is
indeed a low bar.  By having code and data openly deposited, third
parties can later examine the code, and we hope that knowing the code
will be open ensures that authors will not cheat.  It also allows
researchers, potentially with new methods, to look for errors.  This
is more effective than engaging in an arms race on building methods to
detect malicious intent now with closed datasets and code.  This is
analogous to storing blood samples of sport champions today to
possibly detect doping in the future with more sensitive methods
(cf. \cite{everythinghertz97}).  Another comparison that helped us
define the scope of a CODECHECK is that we think of the codechecker as
forensic photographer, capturing details so that an investigator may
later scrutinise them.

\textbf{Who's got time for more peer review?} Agree; codechecking
takes time that could otherwise be used for traditional peer review.
However, a CODECHECK is different from peer review.  First, the
\textbf{technical nature} of a CODECHECK sets clear expectations and
thereby time budget compared to conventional peer review.  For
example, authors are told what to provide and the codechecker can be
told when to stop.  Codecheckers can always directly ask the author
when clarification is required, thereby increasing
\textbf{efficiency}.  Second, the specific skill set of a codechecker
allows for \textbf{different groups} to participate in the review
process.  ECRs might be attracted to learn more about recent methods,
peer review, and reproducibility practices.  Research Software
Engineers  %% (RSE), see \url{https://en.wikipedia.org/wiki/Research_software_engineering})
who might not regularly be involved in writing or reviewing papers
might be interested in increasing their connection with scholarly
practices.  An extra codechecker may simplify the matchmaking an
editor does when identifying suitable reviewers for a submission,
as technical and topical expertise can be provided by different
people.  Third, recall that CODECHECKs should always be publicly
available, unlike peer review reports.  With code and workflows, the
codechecker's feedback may directly impact and improve the author's
work.  The public certificates and contributions provide \textbf{peer
  recognition} for the codechecker.  Fourth, we found that focusing on
workflow mechanics and interacting with the author makes reproductions
\textbf{educational}. It also is a different role and, as such, could
be a welcome option for researchers to give back their time to the
community.

@@

CODECHECK works --- we have created a considerable number of
certificates to demonstrate it.  The creation of certificates and
interactions with authors and editors shaped the principles and the
workflow and also confirmed the approach taken. This result
corroborates findings from similar evaluations of reproducible
computational research in journals and conferences.  CODECHECKs
increase transparency of the checked papers and can contribute to
building trust in scientific results.
% "science earns trust through transparency" > https://twitter.com/briannosek/status/1290240244101521408
The set of shared principles and common name, through recognition
value, will allow researchers to judge the level of scrutiny that
results have faced. CODECHECK requires direct acknowledgement of the
codechecker's contributions, not indirectly via citations of
reproductions or informal credit.

@@

A CODECHECK system, even if temporarily adopted as a sustainable
transition towards more open publication and review practices, can
contribute to increased trust in research outputs. Introducing
CODECHECK should be informed by lessons learned from (introducing)
open peer review \cite{ross-hellauer_guidelines_2019}.  \textbf{Our
conversations with publishers and editors included procedural and
technical implications of openness, be it that open peer review is neither applied
by nor familiar to stakeholders, or that process management software
would not support such open communication. WHAT DOES THIS SENTENCE
MEAN??  JUST DELETE?}

@@

% Local Variables:
% TeX-engine: xetex
% End:
% LocalWords:  preprint reproducibility compendia
